---
title: John Blischak reproducibility and workflowr
author: ''
date: '2018-07-16'
categories:
  - talk
tags:
  - reproducibility
  - reproducible research
  - software
  - computational reproducibility
slug: john-blischak-reproducibility-and-workflowr
header:
  caption: ''
  image: ''
---

Institutional road blocks against full reproducibility - methods sections are insufficient.
But good software practice is helpful for future you, your labmates when you leave.

Lowndes et al (2017)
Peng et al (2011)

## Strategies for computational repro


1. Record computing environment

```{r}
sessionInfo()
```

2. Set the seed! want to ensure exact / direct computational reproducibility before someone can extend your analyses. Ensure point estimates / outputs are the same no matter the computing environment.

3. Organizing into subdirectories

code, data, figures, into distinct subdirectories

using relative paths - whatabout getting around Rmarkdown paths vs. script paths. 

4. Run code in clean environment

But won't remove attached packages:
```{r, eval = FALSE}
rm(list = ls())
```

So better to restart your R session in RStudio.

Best practice: produce final results from command line:

```{bash, eval = FALSE}
Rscript -e 'rmarkdown::render("fit_model.R")'
```



